---
title: "facebook_factcheck"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

need to write a proposal covering these questions:
What is the problem you want to solve?
 The social media world is becoming more and more polarized. One of the early glimpses of this was the 2016 elections. There is a strong element of disinformation throughout social media. This project is based on research done by Buzzzfeed on Facebook leading up to the 2016 presidential elections .That data was acquired over seven week days (September 19 through 23rd and September 26th and 27th, 2016.) 
 The research included 

Who is your client and why do they care about this problem? In other words, what will your client DO or   `DECIDE based on your analysis that they wouldn’t have otherwise?

What data are you going to use for this? How will you acquire this data?

In brief, outline your approach to solving this problem (knowing that this might change later).

What are your deliverables? Typically, this would include code, along with a paper and/or a slide deck.


DATA WRANGLING:

 The data comes from Kaggle. It was submitted by a journalist from the news site Buzzfeed. The data was not very messy. I did have to rename some features.  Account id, post id, and the post url were integers that were changed to character strings. Date Published was a factor that was changed to a date. 
     There were a few NAs that had to be changed in the columns share count, reaction count, and the comment count. We changed the NAs	to the median instead of the mean, because the outliers of all these columns caused each of them to be heavily skewed to the right. 


Possible questions:

How do left, mainstream, and right categories of Facebook pages differ in the stories they share?

Which types of stories receive the most engagement from their Facebook followers? Are videos or links more effective for engagement?

Can you replicate BuzzFeed’s findings that “the least accurate pages generated some of the highest numbers of shares, reactions, and comments on Facebook”?

```{r}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)

fb<-read.csv("file:///C:/Users/John/Documents/R/fact-checking-facebook-politics-pages/facebook-fact-check.csv")

summary(fb)

#check which types of stories recieved more engagement
table(fb$Post.Type)

table(fb$Rating)


```


```{r}
#Renameing features
fb$account_id <- as.character(fb$account_id)
fb$post_id <- as.character(fb$post_id)
fb$Post.URL <- as.character(fb$Post.URL)
fb$Date.Published <- as.Date(fb$Date.Published)

#correcting na
fb[which(is.na(fb$share_count)),"share_count"]<- 87
fb[which(is.na(fb$reaction_count)),"reaction_count"]<- 543
fb[which(is.na(fb$comment_count)),"comment_count"]<- 131


ggplot(fb,aes(x=Category, fill= Rating))+
  geom_bar()

Total_count<- (fb$share_count+ fb$reaction_count+fb$comment_count)

ggplot(fb,aes(x=Rating,y= reaction_count,color= Category))+
  geom_boxplot(coef=3, fun = median)+
  coord_flip()

ggplot(fb, aes(x=Date.Published, y= Total_count))+
  geom_point()

summary(fb)
str(fb)
#View(fb)


```
```{r}
ggplot(fb, aes(x=Page,fill= Rating))+
  geom_bar()+
  coord_flip()

ggplot(fb,aes(Total_count,y= Page,group= reaction_count))+
  stat_boxplot()


ggplot(fb,aes(x=Post.Type,y= reaction_count, col=Rating))+
  geom_point(position = "jitter")

ggplot(data=fb)+
  geom_point(mapping= aes(x=Post.Type, y= Total_count, position= "dodge"))+
  facet_wrap( ~ Rating,nrow = 2)

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
